### 2023-01-06 Manuscript/Draft1.tex
update manuscript and plotting file
manuscript and plotting file update

The Northeast United States continental shelf spans from the Outer Banks of North Carolina to the Gulf of Maine. The region covers over 250,000 km\(^2\) of ocean, extending over 200 km from shore in the largest areas in New England to just 30 km off shore in the southern regions. This ecologically diverse region contains approximately 18,000 vertebrate marine species. Commercial fisheries have been an important part of local economies for centuries. In 2019, New England fisheries produced \$22 billion in sales, which sustained over 200,000 jobs. Maintaining a healthy ecosystem is therefore vital to sustained ecological health and economic prosperity of the region. (NEFMC 2020)

@@

Fish stocks in this highly productive and economically important region are managed by the National Oceanic and Atmospheric Administration's (NOAA) Northeast Fisheries Science Center (NEFSC) in Woods Hole, Massachusetts. Federal scientists assess the health and abundance of each commercial fish stock using fishery-independent bottom trawl survey data that has been collected by NOAA throughout the region since 1963 (cite survey paper{[}{[}, some possible citations are \url{https://repository.library.noaa.gov/view/noaa/4825/noaa_4825_DS1.pdf} and T.S. Azarovitz 1981 A brief historical review of the Woods Hole Laboratory trawl survey time series. W.G. Doubleday, D. Rivard (Eds.), Bottom Trawl Surveys, Can. Spec. Pub. Fish. Aqua. Sci. No.~58 (1981), pp.~62-67{]}{]}). The survey uses a stratified random design where bottom trawl sampling takes place in predefined strata along the eastern continental shelf. The survey has created a rich time series data set with many uses including species-specific habitat identification, analysis of how environmental conditions influence species abundance, and estimating yearly species abundance trends to help inform stock assessments and ultimately quota limits \textbf{just listed a few uses of survey. change/add others?}.

@@

Due to a combination of climate change and shifts in circulation, the Northeast United States continental shelf has experienced rapid warming in recent decades, resulting in a shift in spatial distributions of many species. Since stock assessment models rely on accurate descriptions of population dynamics and contemporary patterns of spatial abundance, there is concern that rapid undocumented changes in spatial distributions of species will bias future stock assessments. The implication of this is that the bottom trawl survey is actually sampling the population during a different life cycle stage than was originally assumed {[}{[}I don't think the concern is about life cycle stage, but rather that the fish catchability or availability in the survey will change, thus changing the relationship between the index and the true population{]}{]}, which can lead to biased stock assessments. We are therefore interested in analyzing the impact of climate change on the accuracy of future stock assessment models as measured by NOAA's ongoing bottom-trawl survey along the East coast.

\begin{itemize}
\tightlist
\item
  say something about our study species? Past abundance history, biological characteristics, others? Maybe better suited in below section {[}{[}I think the specific species can wait for the methods section, just laying the groundwork here in the intro{]}{]}
\end{itemize}

@@

\begin{itemize}
\item
  Fish are changing spatial distribution and have altered life stages (?) because of climate change {[}{[}see note above about catchability and availability{]}{]}
  NYE paper
\item
  Population indexing methods may be becoming biased as a result
\item
  Briefly describe our study to test this
\end{itemize}

@@

We construct spatial models for Yellowtail Flounder, Atlantic Cod, and Haddock on {[}{[}Georges{]}{]} George's Bank, where movement of each species combine static species-specific habitat preferences with temperature preferences. Model dynamics are driven by a time series of temperature gradients that were estimated from data to create simulated data sets for each population where the true biomass is known. Using temperature gradients that repeat each year creates data sets with predictable, repeating spatial patterns, whereas using a temperature gradient that increases on average throughout the simulation leads to spatial distributions that shift over time {[}{[}and mimics the possible future in the region under climate change?{]}{]}. We conducting stratified random sampling on our simulation output to mimic the bottom trawl survey and compare the ability of contemporary indexing methods to track population trends.

@@

We obtained a shapefile for the 15 strata that comprise {[}{[}Georges Bank does not have an apostrophe, I made the same mistake when I started here{]}{]} George's Bank to use as our modeling environment. We discritized the region into a raster with 88 rows and 144 columns. Haddock inhabit all 15 strata in the domain, Cod inhabit 13 strata, and Yellowtail exist in 9 strata. Figure \ref{fig:strata-plot} shows the regions used in our models. {[}{[}can mention that the different strata are based on biological differences among the species, possibly add that there are multiple stocks for all three species and the delineations among the stocks differ by species{]}{]}

@@

The time step for our models is one week. MFS uses a modified two-stage Deriso-Schnute delay difference equation that models the biomass in each cell in our study area (Dolder et al. 2020). Individual terms in the formulation account for growth of mature adults, natural and fishing mortality, and the addition of new recruits. {[}{[}not sure we need to highlight the recruitment model, might be able to just use the table because Bev-Holt formulation is well known in fisheries{]}{]} We chose to represent recruitment in the model using a Beverton-Holt formulation \textbf{cite}. Recruitment is a function of the adult biomass that existed in the previous year and is added to the population incrementally throughout each species' predefined spawning period. Parameter inputs were either obtained from the literature or chosen to produce desired model dynamics. A full list of parameters used in our model can be seen below in Tables \ref{paramsALL} and \ref{tab:paramsSCENARIOS}.

@@

Species-specific habitat preferences were derived using the \emph{lrren} tool from the R package \emph{envi} (Buller 2022) to create a niche model for each species. The \emph{lrren} tool estimates an ecological niche using the relative risk function by relating presence/absence data to two covariate predictors. We used bottom trawl point data in from 2009-2021 as our presence/absence input by using a value of 0 for any tow that failed to catch the given species and weighting a successful catch by the biomass of the given tow {[}{[}I don't think we need to cite the trawl data because we said we were using trawl data above{]}{]} \textbf{cite trawl data?}. We combined data from both the fall and spring surveys to obscure the influence of temperature so that the niche model would instead infer habitat preferences. Depth and mean sediment size were used as our covariate predictors. Estimated depth for the region was obtained from FVCOM (Chen et al. 2006). The mean sediment size raster was interpolated in ArcMap using the natural neighbor interpolation method {[}{[}citing ArcMap should suffice{]}{]} \textbf{cite arcmap or more?} using point data collected by the United States Geologic Survey (USGS) (McMullen et al. n.d.). Since the values in \(Hab^2_{J,s}\) are required to be between 0 and 1, we transform the spatial estimates from \emph{lrren} to fall between these bounds. See Figure \ref{fig:hab-plot1} for a visual representation of this process being applied to Cod. Figure \ref{fig:hab-plot} depicts habitat preferences \(Hab^2_{J,s}\) for each species.

@@

Each species is assumed to have normally distributed temperature preferences (\(N(\mu,\sigma)\)). We assume Yellowtail Flounder's preferences are \(N(8.75,4.25)\), while Haddock and Cod have preferences \(N(9,4)\). We chose these values by combining information in the literature with temperatures recorded in the bottom trawl survey. Weekly estimated temperature data for the region for 2012 was obtained from FVCOM (Chen et al. 2006). We chose 2012 because the data displayed an average temperature pattern that consistently oscillated between maximum and minimum temperature values. This data was also transform to create an oscillating pattern that increases 5 degrees Celsius on average over the duration of the simulation. \textbf{show images of temperature and/or and/or temp videos??? and/or average temperature oscillations?}{[}{[}as we discussed, one plot here comparing average temperatures for constant and increasing scenarios in the paper, one or more animations in supplemental materials{]}{]}

@@

In equation (1), \(Hab^2_{J,s}\) is constant for the duration of the simulation, while \(Tol_{c,s,wk}\) changes each week. Using a temperature gradient that repeats every 52 weeks produces the same spatial preferences in a given week each year, resulting in consistent spatial biomass patterns. Scenarios where the temperature increases over time creates spatial preferences that evolve as the water warms, producing spatial biomass patterns that shift in a given week over the duration of the simulation. Thus, stratified random samples in scenarios with a repeating temperature pattern will have constant survey catchability and availability over time, which may not be true for increasing temperature scenarios due to evolving spatial preferences. {[}{[}may want to mention here or in the discussion that we used repeating 2012 to simplify the simulations and reduce the number of factors explored that would have occurred with different annual cycles{]}{]}

@@

We consider 20 year simulations under three population parameter scenarios for each of our three species- a scenario where parameters result in each population increasing over time, one where the populations are relatively constant over time, and a scenario where the parameter combination results in each population decreasing over time. Each of these three scenarios is paired with a temperature gradient that repeats as well as one that increasing roughly 5 degrees Celsius {[}{[}need some justification for this large change, wanted to be extreme because if the extreme doesn't make a difference, then more reasonable increases wouldn't make a difference{]}{]} over the duration of the 20 year simulation. We therefore simulate a total of 6 scenarios. \textbf{show line plot of population for each?}

@@

After each simulation is complete, we mimic the bottom trawl survey by conducting stratified random sampling in each inhabited strata twice each year. We sample in the same weeks that the Spring and Fall surveys take place and the number of the samples taken in each strata reflect true values. {[}{[}need to make sure that we don't lose sight of random samlping with the following sentences{]}{]} Most strata contain enough cells to sample a unique location in each survey over the duration of the simulation. For smaller strata we must repeat some sample locations. We then use the biomass collected from our samples in contemporary population indexing methods to estimate population trends. Knowing the true population values in our simulations allows us to compare the error calculated from each estimation method.{[}{[}should we mention here or later that we did both with and without measurement error analyses?{]}{]}

@@

We follow the advice given in (Thorson 2019) to build VAST models to estimate biomass on {[}{[}Georges{]}{]} George's Bank using stratified mean samples from our model output. In addition to exploring different link functions and assumed distributions, our VAST model-building process included testing the impact of including spatial and/or spatio-temporal variation in our models, considering varying number of knots in our mesh, and testing different forms of temporal correlation. We also carried out the same process both including covariates in our model as well as running models without covariate information. We considered covariates in the form of dynamic temperature values and/or static habitat values from our population model. When using covariates we ultimately decided to provide the most information to the model by including both covariates for both linear predictors. Since we know the true population values in our models we calculate the absolute error of each VAST estimate to compare between potential settings. Through this process, and in consultation with the VAST package creator, we determined setting that allowed VAST models to converge for all of our scenarios while also providing the lowest absolute error values. Settings for our VAST models can be seen in Table ???.

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.14}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.14}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.14}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.14}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.14}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.18}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.14}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
Species
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Population Trend
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Temperature Scenario
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Strata Included
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Noise in Data
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Covariates (VAST only)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Season
\end{minipage} \\
\midrule
\endhead
Yellowtail & Increasing & Repeating & All strata & No Noise & No Covariates & Spring \\
Cod & Constant & Increasing 5\(^{\circ}\) & Subset & Yes Noise & Temp + Habitat & Fall \\
Haddock & Decreasing & & & & & \\
\bottomrule
\end{longtable}

@@

The goal of our project was to analyze how well contemporary population indexing methods can track population trends under a host of conditions, as depicted in Table COMBOS. Historically, Atlantic Cod has seen significant decline over the last XXX years while Haddock has increased in abundance in recent year {[}{[}can cite the 2022 management track assessments, see \url{https://apps-nefsc.fisheries.noaa.gov/saw/sari.php} for when the document becomes available{]}{]} \textbf{cite}. For this reason we compare indexing estimates using stratified random samples from decreasing population scenarios for Cod (see Table ???) and increasing population scenarios for Haddock (see Table ???). To provide a comprehensive analysis of population indexing methods we consider all possible scenario combinations for Yellowtail Flounder (see Table ???).

@@

{[}{[} some things that I think we'll need to address, in addition to those mentioned above are (in no particular order):
number of simulations done for each scenario - time required to do them and analyses
pros and cons of using VAST with covariates in a changing environment
implications for current practice of using stratified mean (not horrible, but might be able to do better with model-based using covariates if enviro changing, note we don't want to say stratified mean should not be used because that will draw a lot of tomato throwing)
future work - some of the ideas we've discussed already could be mentioned
limitations to our study{]}{]}

@@

{[}{[}we should thank Jim (obviously), but also those who helped you with the habitat data and others? also should note that funding provided by Climate-Groundfish source (I'll dig up the official name of the funding source){]}{]}

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-envi}{}}%
Buller, I.D. 2022. Envi: Environmental interpolation using spatial kernel density estimation. The Comprehensive R Archive Network. doi:\href{https://doi.org/10.5281/zenodo.5347826}{10.5281/zenodo.5347826}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-nefmc20}{}}%
NEFMC. 2020. {Northeast skate complex fishery management plan. Amendment 5 discussion document}. {Available online at https://www.nefmc.org/library/amendment-5-3}.

@@

\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
fac774b6b2cab2991e9875992ad8ca7eb3de1326
\end{quote}
\end{quote}
\end{quote}
\end{quote}
\end{quote}
\end{quote}
\end{quote}

>>>>

The Northeast United States continental shelf spans from the Outer Banks of North Carolina to the Gulf of Maine. The region covers over 250,000 km\(^2\) of ocean, extending over 200 km from shore in the largest areas in New England to just 30 km off shore in the southern regions. This ecologically diverse region contains approximately 18,000 vertebrate marine species. Commercial fisheries have been an important part of local economies for centuries. In 2019, New England fisheries produced \$22 billion in sales, which sustained over 200,000 jobs. Maintaining a healthy ecosystem is therefore vital to sustained ecological health and economic prosperity of the region.

@@

Fish stocks in this highly productive and economically important region are managed by the National Oceanic and Atmospheric Administration's (NOAA) Northeast Fisheries Science Center (NEFSC) in Woods Hole, Massachusetts. Federal scientists assess the health and abundance of each commercial fish stock using fishery-independent bottom trawl survey data that has been collected by NOAA throughout the region since 1963 (Azarovitz 1981; Politis et al. 2014). The survey uses a stratified random design where bottom trawl sampling takes place in predefined strata along the eastern continental shelf. The survey has created a rich time series data set with many uses including species-specific habitat identification, analysis of how environmental conditions influence species abundance, and estimating yearly species abundance trends to help inform stock assessments and ultimately quota limits \textbf{just listed a few uses of survey. change/add others?}.

@@

Due to a combination of climate change and shifts in circulation, the Northeast United States continental shelf has experienced rapid warming in recent decades, resulting in a shift in spatial distributions of many species. Since stock assessment models rely on accurate descriptions of population dynamics and contemporary patterns of spatial abundance, there is concern that rapid undocumented changes in spatial distributions of species will bias future stock assessments. More specifically, as fish populations shift their distributions over time, catchability and/or availability in the survey will change, altering the relationship between the index and the true population. We are therefore interested in analyzing the impact of climate change on the accuracy of future stock assessment models as measured by NOAA's ongoing bottom-trawl survey along the East coast.

@@

\begin{itemize}
\tightlist
\item
  Briefly describe our study to test this
\end{itemize}

@@

We construct spatial models for Yellowtail Flounder, Atlantic Cod, and Haddock on Georges Bank, where movement of each species combine static species-specific habitat preferences with temperature preferences. Model dynamics are driven by a time series of temperature gradients that were estimated from data to create simulated data sets for each population where the true biomass is known. Using temperature gradients that repeat each year creates data sets with predictable, repeating spatial patterns, whereas using a temperature gradient that increases on average throughout the simulation leads to spatial distributions that shift over time, approximating the future distribution in the region under climate change. We conducting stratified random sampling on our simulation output to mimic the bottom trawl survey and compare the ability of contemporary indexing methods to track population trends.

@@

We obtained a shapefile for the 15 strata that comprise Georges Bank, where strata were partitioned based primarily on depth and secondarily by latitude (Politis et al. 2014). The region was discritized into a raster with 88 rows and 144 columns to use as our modeling environment. A fish stock is considered to be a subpopulation of a species that has similar intrinsic parameters. Each of the species being modeled has multiple distinct stocks along the Atlantic coast resulting from local environmental conditions. Biological differences between species results in each stock inhabiting a different number of strata on Georges Bank. Haddock inhabit all 15 strata in the domain, Cod inhabit 13 strata, and Yellowtail exist in 9 strata. Figure \ref{fig:strata-plot} shows the regions used in our models.

@@

The time step for our models is one week. MFS uses a modified two-stage Deriso-Schnute delay difference equation that models the biomass in each cell in our study area (Dolder et al. 2020). Individual terms in the formulation account for growth of mature adults, natural and fishing mortality, and the addition of new recruits. {[}{[}not sure we need to highlight the recruitment model, might be able to just use the table because Bev-Holt formulation is well known in fisheries{]}{]}\textbf{BL: MFS has 2 recruitment options and I edited the package to use previous years biomass} We chose to represent recruitment in the model using a Beverton-Holt formulation \textbf{cite}. Recruitment is a function of the adult biomass that existed in the previous year and is added to the population incrementally throughout each species' predefined spawning period. Parameter inputs were either obtained from the literature or chosen to produce desired model dynamics. A full list of parameters used in our model can be seen below in Tables \ref{paramsALL} and \ref{tab:paramsSCENARIOS}.

@@

Species-specific habitat preferences were derived using the \emph{lrren} tool from the R package \emph{envi} (Buller 2022) to create a niche model for each species. The \emph{lrren} tool estimates an ecological niche using the relative risk function by relating presence/absence data to two covariate predictors. We used bottom trawl point data in from 2009-2021 as our presence/absence input by using a value of 0 for any tow that failed to catch the given species and weighting a successful catch by the biomass of the given tow {[}{[}I don't think we need to cite the trawl data because we said we were using trawl data above{]}{]} \textbf{BL: Does trawl data ever get cited? It was not cited above. cite trawl data?}. We combined data from both the fall and spring surveys to obscure the influence of temperature so that the niche model would instead infer habitat preferences. Depth and mean sediment size were used as our covariate predictors. Estimated depth for the region was obtained from FVCOM (Chen et al. 2006). The mean sediment size raster was interpolated in ArcMap using the natural neighbor interpolation method \textbf{cite arcmap} using point data collected by the United States Geologic Survey (USGS) (McMullen et al. n.d.). Since the values in \(Hab^2_{J,s}\) are required to be between 0 and 1, we transform the spatial estimates from \emph{lrren} to fall between these bounds. See Figure \ref{fig:hab-plot1} for a visual representation of this process being applied to Cod. Figure \ref{fig:hab-plot} depicts habitat preferences \(Hab^2_{J,s}\) for each species.

@@

Each species is assumed to have normally distributed temperature preferences (\(N(\mu,\sigma)\)). We assume Yellowtail Flounder's preferences are \(N(8.75,4.25)\), while Haddock and Cod have preferences \(N(9,4)\). We chose these values by combining information in the literature with temperatures recorded in the bottom trawl survey. Weekly estimated temperature data for the region for 2012 was obtained from FVCOM (Chen et al. 2006). The 2012 data was repeated yearly in the constant temperature scenario because displayed an average temperature pattern that consistently oscillated between maximum and minimum temperature values. We chose to repeat estimates for a single year rather than use data for consecutive years to reduce the number of factors impacting model dynamics while still incorporating real data. The 2012 temperature data was also transform to create an oscillating pattern that increases 5 degrees Celsius on average over the duration of the simulation. We chose a 5 degree increase over a 20 year simulation to allow temperature change to have a meaningful impact on dynamics while remaining within reasonable computational limits in terms of the length of the simulation. Figure \ref{fig:temp-scenarios} depicts mean trends for the temperature scenarios used in our models. \textbf{dont forget to include gif in final submission}

{\centering \includegraphics[width=0.95\linewidth]{Images/TempScenarios}

\caption{Mean trends of temperature data used in our model.}\label{fig:temp-scenarios}
\end{figure}

@@

In equation (1), \(Hab^2_{J,s}\) is constant for the duration of the simulation, while \(Tol_{c,s,wk}\) changes each week. Using a temperature gradient that repeats every 52 weeks produces the same spatial preferences in a given week each year, resulting in consistent spatial biomass patterns. Scenarios where the temperature increases over time creates spatial preferences that evolve as the water warms, producing spatial biomass patterns that shift in a given week over the duration of the simulation. Thus, stratified random samples in scenarios with a repeating temperature pattern will have constant survey catchability and availability over time, which may not be true for increasing temperature scenarios due to evolving spatial preferences. {[}{[}\textbf{BL: included in above paragraph} may want to mention here or in the discussion that we used repeating 2012 to simplify the simulations and reduce the number of factors explored that would have occurred with different annual cycles{]}{]}

@@

We consider 20 year simulations under three population parameter scenarios for each of our three species- a scenario where parameters result in each population increasing over time, one where the populations are relatively constant over time, and a scenario where the parameter combination results in each population decreasing over time. Each of these three scenarios is paired with a temperature gradient that repeats as well as one that increasing roughly 5 degrees Celsius {[}{[}\textbf{BL: included in above paragraph} need some justification for this large change, wanted to be extreme because if the extreme doesn't make a difference, then more reasonable increases wouldn't make a difference{]}{]} over the duration of the 20 year simulation. We therefore simulate 6 scenarios for each population.

@@

After each simulation is complete, we mimic the bottom trawl survey by conducting stratified random sampling in each inhabited strata twice each year. We sample each strata in the same weeks in which the Spring and Fall surveys take place and the number of the samples taken reflect true target values for each strata. Most strata contain enough cells that we were able to generate a list of sampling locations without replacement, allowing us to sample a unique random cell in each survey over the duration of the simulation. For strata that contain small numbers of cells, we generate a random sampling locations with replacement of cells previously sampled. We then use the biomass collected from our samples in contemporary population indexing methods to estimate population trends. Knowing the true population values in our simulations allows us to compare the error calculated from each estimation method.
\textbf{BL: added both to above paragraph}
{[}{[}should we mention here or later that we did both with and without measurement error analyses?{]}{]}
{[}{[}need to make sure that we don't lose sight of random sampling with the following sentences{]}{]}

@@

We follow the advice given in (Thorson 2019) to build VAST models to estimate biomass on Georges Bank using stratified mean samples from our model output. In addition to exploring different link functions and assumed distributions, our VAST model-building process included testing the impact of including spatial and/or spatio-temporal variation in our models, considering varying number of knots in our mesh, and testing different forms of temporal correlation. We also carried out the same process both including covariates in our model as well as running models without covariate information. We considered covariates in the form of dynamic temperature values and/or static habitat values from our population model. When using covariates we ultimately decided to provide the most information to the model by including both covariates for both linear predictors. Since we know the true population values in our models we calculate the absolute error of each VAST estimate to compare between potential settings. Through this process, and in consultation with the VAST package creator, we determined setting that allowed VAST models to converge for all of our scenarios while also providing the lowest absolute error values. Settings for our VAST models can be seen in Table ???.

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.14}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.14}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.19}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.14}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.14}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.14}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.10}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
Species
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Population Trend
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Temperature Scenario
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Strata Included
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Noise Added
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Covariates (VAST)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Season
\end{minipage} \\
\midrule
\endhead
Yellowtail & Increasing & Repeating & All strata & No & No Covariates & Spring \\
Cod & Constant & Increasing 5\(^{\circ}\) & Subset & Yes & Temp + Habitat & Fall \\
Haddock & Decreasing & & & & & \\
\bottomrule
\end{longtable}

@@

The goal of our project was to analyze how well contemporary population indexing methods can track population trends under a host of conditions, as depicted in Table COMBOS. Historically, Atlantic Cod has seen significant decline over the last XXX years while Haddock has increased in abundance in recent year {[}{[}can cite the 2022 management track assessments, see \url{https://apps-nefsc.fisheries.noaa.gov/saw/sari.php} for when the document becomes available{]}{]} \textbf{cite}. For this reason we compare indexing estimates using stratified random samples from decreasing population scenarios for Cod (see Table ???) and increasing population scenarios for Haddock (see Table ???). To provide a comprehensive analysis of population indexing methods we consider all possible scenario combinations for Yellowtail Flounder (see Table ???). The specific population trends used in our analyses can be see in Figure \ref{fig:pop-scenarios}

{\centering \includegraphics[width=0.99\linewidth]{Images/Population_scenarios}

\caption{True population trends used in indexing analyses. Spring biomass plots are shown with fall values being very similar.}\label{fig:pop-scenarios}
\end{figure}

@@

{[}{[} some things that I think we'll need to address, in addition to those mentioned above are (in no particular order):
number of simulations done for each scenario - time required to do them and analyses
pros and cons of using VAST with covariates in a changing environment
implications for current practice of using stratified mean (not horrible, but might be able to do better with model-based using covariates if enviro changing, note we don't want to say stratified mean should not be used because that will draw a lot of tomato throwing)
future work - some of the ideas we've discussed already could be mentioned
limitations to our study
perhaps for discussion, but we should mention somewhere that MFS only tracks biomass not age structure, the latter is important for stock assessments but not considered in this paper{]}{]}

@@

{[}{[}we should thank Jim (obviously), but also those who helped you with the habitat data (robyn, david Chev.) and others? also should note that funding provided by Climate-Groundfish source (I'll dig up the official name of the funding source){]}{]}

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-azarovitz1981brief}{}}%
Azarovitz, T. 1981. A brief historical review of the woods hole laboratory trawl survey time series. Bottom trawl surveys.

\leavevmode\vadjust pre{\hypertarget{ref-envi}{}}%
Buller, I.D. 2022. Envi: Environmental interpolation using spatial kernel density estimation. The Comprehensive R Archive Network. doi:\href{https://doi.org/10.5281/zenodo.5347826}{10.5281/zenodo.5347826}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-politis2014northeast}{}}%
Politis, P.J., Galbraith, J.K., Kostovick, P., and Brown, R.W. 2014. Northeast fisheries science center bottom trawl survey protocols for the NOAA ship henry b. bigelow.

@@
